---
title: "Analysis report for `r snakemake@config[['library']][['name']]`"
author: "Quio"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: hide # or show
    toc: TRUE
    toc_float:
      collapsed: TRUE
    toc_depth: 5
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F,
                      warning = F)
```
```{r,include=F, eval=T}
# save.image("html_report_image.RData")
```

```{r, libraries, message=F, error=F}
library(tidyverse)
library(patchwork) # to arrange plots side by side, in grid, etc
library(rjson)
library(ggtext) # To fit long text in ggplot
library(openxlsx) # To write xls files (with format and everything)
library(readxl) # To read exel files (I think this can also be done with the package above)
library(ggrastr) 
```
```{r, libraries_not_in_use,include=F, message=F, error=F}
# library(clipr) # to copy text to the clipboard
# library(tidygenomics) 
# library(Hmisc) # to divide the vector in equal parts using cut2()
# library(biomaRt)# To get gene names out of refseq accessions
# library(ggrepel) # To put names on plots
# library(ggfortify) # To plot PCAs with ggplot
# library(factoextra) # To easily get info from PCA analysis
# library(ViSEAGO) # To analyze GO terms, although I don't think I am going to use it.
# library(qpcR) # To bind vectors of different length using qpcR:::cbind.na
# library(eulerr) # To draw proportional venn diagrams (need to have the numbers though)
# library(R.utils) # To count number of lines of a file
# <br/><br/>    ---> That is to insert a break in the page so that the text don't wrap the images when they don't occupy the whole width
```
# General information  
`r snakemake@config[['library']][['general_info']]`

# Load data  

## Aggregate profiles   
```{r, load_agg_profs}
# for debugging
  # setwd("..")

# Set paramter
b6_profile <- !is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs) |
                  !is.null(snakemake@input$mm10_top_5000_ag_profs)

# Load B6
if (!is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs)) {
  load(snakemake@input[["mm10_top5000_asmtric_auto_XnonPAR_ag_profs"]])
} else if (!is.null(snakemake@input$mm10_top_5000_ag_profs) ) {
  load(snakemake@input$mm10_top_5000_ag_profs)
  # Change name of object to match the name of same object when 'agg_profiles' is loaded
  aggregate_profiles <- list()
  aggregate_profiles[["top_5000_plus_minus_2000"]] <- 
    final_table 
}

# Load B6xCAST
if (!is.null(snakemake@input$B6xCAST_top_5000_pm_1000bp)) {
  load(snakemake@input$B6xCAST_top_5000_pm_1000bp)
  B6xCAST_top_5000_pm_1000bp <- final_table
  
  load(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand)
  B6xCAST_PRDM9_assymetric_hs_invading_strand <- final_table
  
  load(snakemake@input$B6xCAST_PRDM9_assymetric_hs_receiving_strand)
  B6xCAST_PRDM9_assymetric_hs_receiving_strand <- final_table
  rm(final_table)
  
  B6xCAST_PRDM9_assymetric_hs <- 
    bind_rows(B6xCAST_PRDM9_assymetric_hs_invading_strand, 
              B6xCAST_PRDM9_assymetric_hs_receiving_strand)
  
  load(snakemake@input$B6xCAST_PRDM9_assymetric_hs_mm10_aligned)
  B6xCAST_PRDM9_assymetric_hs_mm10_aligned <- final_table
  rm(final_table)
  
  rm(B6xCAST_PRDM9_assymetric_hs_receiving_strand,
     B6xCAST_PRDM9_assymetric_hs_invading_strand)
}
```

## Peaks  
```{r}
# Peak summary
  # Peaks found at "Results/{genomes_not_fused}/Analysis/Peaks_summary.tsv"
if (is.null(snakemake@input[["peaks_summary"]])) {
  "No peaks were asked for"
} else{
  peaks <- read_tsv(snakemake@input[["peaks_summary"]])
  peaks_sorted <-
    peaks %>%
    arrange(desc(broad_all)) %>%
    select(any_of(
      c(
        "sample" = "sample_name",
        "broad_all",
        "broad_hs",
        "% brd peaks at HS",
        "narrow_all",
        "narrow_hs",
        "% nrw peaks at HS"
      )
    ))
}
```

## Number of reads in fastq and in final bam  
How may reads were in the fastq used for alignment (after fastp filtering), and how many primary reads on the final filtered BAM file.  
```{r, get_number_of_reads}
# Get fastp summary json files
files_fastp <-  
  snakemake@input[["fastp"]]

files_fastp_names <-
  snakemake@input[["fastp"]] %>% 
  basename %>% 
  str_remove("\\.(PE|SE)\\.fastp\\.json")

# Get number of reads in FASTQ file from fastp json files
FASTQ_reads <-
  map_dbl(files_fastp,
          ~fromJSON(file = .) %>%
            .[['read1_after_filtering']] %>%
            .[['total_reads']]/1000000) %>% 
  round(2) %>% format(nsmall = 2)

results_reads_filtered_FASTQs <- 
  tibble("sample" = files_fastp_names,
         "# R1 in filtered FASTQ" = FASTQ_reads)

# Final bam file
  # Get the info from the field "read1" (since they should all be properly paired).
get_read1_number <- 
  function(flagstat_text){
    read1 <- readLines(flagstat_text) %>% 
                .[str_detect(.,"read1")] %>%
                str_extract("^[0-9]+") %>% 
                as.double()
    if (read1 == 0) { # When SR, read1 and read2 are == 0 in flagstat
      read1 <- readLines(flagstat_text) %>% 
                .[str_detect(.,"in total")] %>%
                str_extract("^[0-9]+") %>% 
                as.double()
    read1
    }
    read1 <-   
      read1 %>%
      `/` (1000000) %>% # that's the way to do this... weird...
      round(1) %>% 
      format(nsmall = 2)
}

bam_read_file_names <- 
  snakemake@input[["bamfiles_reads"]] %>% 
  basename %>% 
  str_remove(".flagstat.txt")

filtered_BAM_reads <-
  map_chr(snakemake@input[["bamfiles_reads"]],
          get_read1_number)

results_reads_processed_bams <- 
  tibble("sample" = bam_read_file_names,
         "# R1 in filtered BAM" = filtered_BAM_reads)
     
# Put everything together
results_reads <- 
  full_join(results_reads_filtered_FASTQs,
            results_reads_processed_bams)
```


# Quality control  
For a detailed quality control analysis look at the multiqc report (Results/genome/Qctrl/Multiqc...)

# Number of reads and peaks  
```{r}
if (is.null(snakemake@input[["peaks_summary"]])){
  DT::datatable(
    results_reads,
    filter = "top")
} else {
  DT::datatable(
    full_join(peaks_sorted,results_reads),
    filter = "top")
}
```

# Plots  

## Hotspots  

### B6 hotspots  

#### Top 5000 Hotspots  

```{r, include=!b6_profile}
"Aggregate profiles were not requested for any of the samples"
```

```{r, eval = b6_profile, include = b6_profile}
prots <- levels(aggregate_profiles$top_5000_plus_minus_2000$Protein)
names(prots) <- prots # So that name of each protein is printed (in the next loop) and can be searched in the web browser as text

lapply(prots,
  function(i) {
    strands_n <- aggregate_profiles$top_5000_plus_minus_2000$Strand %>% unique %>% length
    a <- ggplot(data = filter(aggregate_profiles$top_5000_plus_minus_2000 ,
                        Protein == i,
                        Strand == "Both_strands")) +
    geom_line(mapping = aes(x = Coordinates,
                            y = Mean_Max_Normalized_Coverage)) +
      theme(legend.position="top")
   if (strands_n > 1){
     b <- ggplot(data = filter(aggregate_profiles$top_5000_plus_minus_2000,
                        Protein == i,
                        Strand != "Both_strands")) +
    geom_line(mapping = aes(x = Coordinates,
                            y = Mean_Max_Normalized_Coverage,
                            color = Strand)) +
     theme(legend.position="top")
     rasterize(a|b)
   } else {
     rasterize(a)
   }
  }
  )
```

#### X-nonPAR vs autosomal  
This analysis and the ones following (at hotspots) are only shown for single stranded profiles.  
Only one strand is shown at a set of 547 X-non-PAR hotspots and their (SPO11-intensity) matched autosomal counterparts.  
```{r, include = is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs)}
"There are no single strande profiles in this dataset and hence X-nonPAR hotspot's aggregate profile is not performed. If you want to see this report you need to set 'get_single_strand' and 'top5000_HS_heatmap' as true"
```
```{r, eval=!is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs), include=!is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs)}
# Get name of one of the strands (this changes whether the sequencing was SR or PE -). This is to filter by that strand and show only one strand.
ss_prots <- levels(aggregate_profiles$xNonPar_autosomal$Protein)
names(ss_prots) <- ss_prots

lapply(ss_prots,
       function(i) {
         normalized <-
           ggplot(
             data = filter(
               aggregate_profiles$xNonPar_autosomal,
               Protein == i,
               Strand != "Both_strands"
             ),
             aes(x = Coordinates,
                 y = Mean_Max_Normalized_Coverage,
                 color = Region)
           ) +
           geom_line() +
           theme(legend.position = "top",
                 plot.title = element_textbox_simple(hjust = 0,
                                                     size = 10)) +
           facet_wrap( ~ Strand)
         
         not_normalized <-
           ggplot(
             data = filter(
               aggregate_profiles$xNonPar_autosomal,
               Protein == i,
               Strand != "Both_strands"
             ),
             aes(x = Coordinates,
                 y = Average_signal,
                 color = Region)
           ) +
           geom_line() +
           theme(legend.position = "top")
         rasterize(normalized | not_normalized) +
           facet_wrap( ~ Strand)
       })
```

#### Asymetric Hotspots  
The plots on the left are with each side, strong and weak, at different sides of the DSB (as they are). On the plot on the right, both sides are plotted on the same side of the DSB so that they can be directly compared.  
```{r, include=is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs)}
"There are no single strand profiles in this dataset and hence asymetric hotspot's aggregate profile is not performed. If you want to see this report you need to set 'get_single_strand' and 'top5000_HS_heatmap' as true"
```

```{r, eval=!is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs), include=!is.null(snakemake@input$mm10_top5000_asmtric_auto_XnonPAR_ag_profs)}
lapply(ss_prots,
  function(i) {
   on_diff_sides <- 
     ggplot(data =
              filter(aggregate_profiles$asymetric,
                                    Protein == i)) +
     geom_line(mapping = aes(x = Coordinates, 
                             y = Average_signal,
                             color = Strength)) +
     labs(title = 
            str_c(
              "Strengths in different sides of DSB")) +
     theme(legend.position="top")
   
   on_same_sides <-  
     ggplot(data = filter(aggregate_profiles$asymetric_flipped,
                          Protein == i)) +
     geom_line(mapping = aes(x = Coordinates, 
                             y = Average_signal,
                             color = Strength)) +
     labs(title = str_c("Strengths on same side of DSB")) +
     theme(legend.position="top")
   
   plots <- on_diff_sides | on_same_sides 
   rasterize(plots + plot_annotation(title = "Left-right assymetric hotspots"))
  })
```

### B6xCAST hotspots  

#### Top 5000 hotspots  
Hotspot list was taken from DMC1 ChIP-seq done with kinetic enrichement by [Smagulova *et al*](http://genesdev.cshlp.org/content/30/3/266) on B6xCAST mix.  
```{r, include=is.null(snakemake@input$B6xCAST_top_5000_pm_1000bp)}
"No B6xCAST experiments were done on this library"
```
```{r, eval=!is.null(snakemake@input$B6xCAST_top_5000_pm_1000bp)}
prots <- levels(B6xCAST_top_5000_pm_1000bp$Protein)
names(prots) <- prots

lapply(prots,
  function(i) {
    strands_n <- B6xCAST_top_5000_pm_1000bp$Strand %>% nlevels
    a <- ggplot(data = filter(B6xCAST_top_5000_pm_1000bp ,
                        Protein == i,
                        Strand == "Both_strands")) +
    geom_line(mapping = aes(x = Coordinates,
                            y = Mean_Max_Normalized_Coverage)) +
      theme(legend.position="top")
   if (strands_n > 1){
     b <- ggplot(data = filter(B6xCAST_top_5000_pm_1000bp,
                        Protein == i,
                        Strand != "Both_strands")) +
    geom_line(mapping = aes(x = Coordinates,
                            y = Mean_Max_Normalized_Coverage,
                            color = Strand)) +
     theme(legend.position="top")
     rasterize((a | b) + plot_annotation(title="B6xCAST top 5000 hotspots"))
   } else {
     rasterize(a + plot_annotation(title="B6xCAST top 5000 hotspots"))
   }
  }
  )
```

#### PRDM9-binding assymetric hotspots    
The list of hotpots used here corresponds to a subset of B6xCAST mix hotspots in which PRDM9 binding is asymmetric between CAST and B6 DNA. Knowing which strand  binds PRDM9 with highest affinity allows us to determing which strand is going to get cut (preferentially), and which one is going to serve as template for repair.  
We refer to the strand that gets preferentially cut as the **invading strand**, and the one that works as template as the **receiving strand**.   
```{r, include=is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand)}
"No B6xCAST experiments were done on this library"
```

##### mm10 aligned reads  
Here are the aggregate profiles of mm10-aligned reads on these set of PRDM9-asymmetric hotspots. On these plots there is no discrimination between 'invading' or 'receiving' DNA.  
```{r, include=!is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand), eval=!is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand)}
prots <- levels(B6xCAST_PRDM9_assymetric_hs_mm10_aligned$Protein)
names(prots) <- prots

map(prots,
    function (i) {
      a <-
        ggplot(
          data = filter(
            B6xCAST_PRDM9_assymetric_hs_mm10_aligned ,
            Protein == i,
            Strand == "Both_strands"
          )
        ) +
        geom_line(mapping = aes(x = Coordinates,
                                y = Mean_Max_Normalized_Coverage)) +
        theme(legend.position = "top")
      
      b <-
        ggplot(
          data = filter(
            B6xCAST_PRDM9_assymetric_hs_mm10_aligned,
            Protein == i,
            Strand != "Both_strands"
          )
        ) +
        geom_line(mapping = aes(x = Coordinates,
                                y = Mean_Max_Normalized_Coverage,
                                color = Strand)) +
        theme(legend.position = "top")
      
      rasterize( a |b)
    })
```


##### Invading or receiving  
Here are the aggregate profiles for both strands of DNA (watson or crick) on either 'invading' or 'receiving' DNA (B6 or CAST depending on the hostpot).  
```{r, include=!is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand), eval=!is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand)}

prots <- levels(B6xCAST_top_5000_pm_1000bp$Protein)
names(prots) <- prots

map(prots,
    ~ rasterize(
      ggplot(
        B6xCAST_PRDM9_assymetric_hs %>% filter(Strand != "Both_strands",
                                               Protein == .x),
        aes(x = Coordinates,
            y = Average_signal,
            color = Strand)
      ) +
        geom_line() +
        facet_wrap(~ Role, scales = "free") +
        theme(legend.position = "top")
    ))
```

##### Invading vs receiving  
Here is the same information as in previous section but both invading and receiving of one strand (crick or watson) are plot together for comparison.  
```{r, include=!is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand), eval=!is.null(snakemake@input$B6xCAST_PRDM9_assymetric_hs_invading_strand)}

map(prots,
    ~ rasterize(
      ggplot(
        B6xCAST_PRDM9_assymetric_hs %>% filter(Protein == .x,
                                               Strand != "Both_strands"),
        aes(x = Coordinates,
            y = Average_signal,
            color = Role)
      ) +
        geom_line() +
        facet_wrap(~ Strand) +
        theme(legend.position = "top")
    ))
```

# Export data  
Ignore this part, it will soon be removed from the report.  

## Excel file for NGS_files.xlsx
```{r}
excel_col_names <- "experiment_name	protein	species	library	number_broad_peaks	number_narrow_peaks	%_broad_peaks_in_HS_+-_2000bp	%_narrow_peaks_in_HS_+-_2000bp	ra_synchronized	broad_meiotic_stage	specific_meiotic_stage	unique_q_filtered_reads[M]	sequencing_type	sequencing_length	sequencing_equipment	library_prep_method	Storage	Container	Full_path	file_name	narrow_peaks_in_HS_+-_2000bp	broad_peaks_in_HS_+-_2000bp" %>% str_split("\t") %>% unlist
```
```{r}
if (is.null(snakemake@input[["peaks_summary"]])){
  table <- results_reads %>% 
    mutate("broad_all" = "-",
           "narrow_all" = "-",
           "% brd peaks at HS" = "-",
           "% nrw peaks at HS" = "-",
           "narrow_hs" = "-",
           "broad_hs" = "-")
} else {
  table <- left_join(peaks_sorted,results_reads)
}

table2 <- 
  table %>% 
  mutate(protein = "-",
         species = "-",
         library = "-",
         ra_synchronized = "-",
         broad_meiotic_stage = "-",
         specific_meiotic_stage = "-",
         sequencing_type = "-",
         sequencing_length = "-",
         sequencing_equipment = "-",
         library_prep_method = "-",
         Storage = "-",
         Container = "-",
         Full_path = "-",
         file_name = "-") %>% 
  rename(any_of(c("experiment_name" = "sample",
         "number_broad_peaks" = "broad_all",
         "number_narrow_peaks" = "narrow_all",
         "%_broad_peaks_in_HS_+-_2000bp" = "% brd peaks at HS",
         "%_narrow_peaks_in_HS_+-_2000bp" = "% nrw peaks at HS",
         "unique_q_filtered_reads[M]" = "# R1 in filtered BAM",
         "narrow_peaks_in_HS_+-_2000bp" = "narrow_hs",
         "broad_peaks_in_HS_+-_2000bp" = "broad_hs"))) %>% 
  dplyr::select(any_of(excel_col_names))

# Create header style for excel file
hs <- createStyle(
  textDecoration = "BOLD", fontColour = "#FFFFFF", fontSize = 12,
  fontName = "Arial Narrow", fgFill = "#4F80BD"
)

# Write excel file
write.xlsx(table2,
           file = str_c(snakemake@output$xlsx_table),
           colNames = TRUE,
           borders = "all",
           firstRow = TRUE,
           colWidths = "auto",
           headerStyle = hs,
           zoom = 150,
           overwrite = T)
```

# Session Info  
```{r session_info}
utils::sessionInfo()
```
